Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2023-02-03 analysis/paper/paper.qmd
Finished editing results

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point(size = 3) +
    labs(x = "Weight (g)", 
       y = "TCSA",
       colour='Raw material') +
  geom_smooth(method=lm, se=FALSE, size = 1)+
  theme_bw(base_size = 10) +
  theme(
  legend.title = element_text(size = 6),
  legend.text = element_text(size = 5),
  legend.position = c(.8,.3),
  legend.key.height = unit(4.5, 'mm'),
  legend.key.width = unit(2, 'mm'))

@@

@@

We computed artifact TCSA values for `r n_sites` sites that contain more than 5 stemmed points to observe variation between sites (@fig-tcsa-per-sites) Sites with fewer than 5 stemmed points were grouped under the category of "Other". Among the sites in our sample, Suyanggae has the most stemmed points and the widest range of TCSA values. This suggests that people made stemmed points for accomplishing a variety of tasks at Suyanggae site. Nosanri, Sibuk, and Wolpyeng have a narrower range of lower values, which indicates less diverse likely uses for stemmed points. Yongsujaeul shows a narrower range but higher TCSA value. Yongsandong has the second most stemmed points and TCSA shows a similar distribution to Yongsakul.

@@

In order to explore spatial patterns and the relationship between the role of stemmed points and the environment, we located individual assemblages on maps that depict eco-regional zones. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of vegetation map or Central inland (CI) (@fig-gis-tcsa-sites panel A & D) of eco-region map. We observed that there are statistically significant differences in the distributions of TCSA values between some zones (Vegetation zones: `r eco_lgm_test_pretty`, Ecoregion zones: `r for_lgm_test_pretty`). The Tukey's HSD results show that TCSA values from the South Temperature Zone (ST) zone are significantly different from the other two vegetation zones (@fig-gis-tcsa-sites panel C). Looking into this further, we see that there are only six artifacts from the ST zone, with only three of these having TSCA values above 175. The other three have TCSA values lower than 100, similar to the other zones (@fig-gis-tcsa-sites panel B). Given the small number of artifacts with extreme TCSA values in the ST zone, we hesitate to conclude this is an archaeologically significant pattern of different TCSA values. The Tukey HSD pairwise differences show no pairs of eco-regions with significant differences in TCSA values. Overall, our results indicate no clear patterns in the distribution of TCSA values across the vegetation and eco-regional zones.

>>>>

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point(size = 1.5) +
    labs(x = "Weight (g)", 
       y = "TCSA",
       colour='Raw material') +
  geom_smooth(method=lm, se=FALSE, size = .7)+
  theme_bw(base_size = 10) +
  theme(
  legend.title = element_text(size = 6.3),
  legend.text = element_text(size = 5.3),
  legend.position = c(.8,.3),
  legend.key.height = unit(4.8, 'mm'),
  legend.key.width = unit(2.5, 'mm'))

@@

We computed artifact TCSA values for `r n_sites` sites that contain more than 5 stemmed points to observe variation between sites (@fig-tcsa-per-sites). Sites with fewer than 5 stemmed points were grouped under the category of "Other". Among the sites in our sample, Suyanggae has the most stemmed points and the widest range of TCSA values. This suggests that people made stemmed points for accomplishing a variety of tasks at Suyanggae site. Nosanri, Sibuk, and Wolpyeng have a narrower range of lower values, which indicates less diverse likely uses for stemmed points. Yongsujaeul shows a narrower range but higher TCSA value. Yongsandong has the second most stemmed points and shows roughly two different clusters of TCSA values.

@@

In order to explore spatial patterns and the relationship between the role of stemmed points and the environment, we located individual assemblages on maps that depict eco-regional zones. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of vegetation map or Central inland (CI) (@fig-gis-tcsa-sites panel A & D) of eco-region map. We observed that there are statistically significant differences in the distributions of TCSA values between some zones (Vegetation zones: `r eco_lgm_test_pretty`, Ecoregion zones: `r for_lgm_test_pretty`). The Tukey's HSD results show that TCSA values from the South Temperature Zone (ST) zone are significantly different from the other two vegetation zones (@fig-gis-tcsa-sites panel C). Looking into this further, we see that there are only six artifacts from the ST zone, with only three of these having TSCA values above 175. The other three have TCSA values lower than 100, similar to the other zones (@fig-gis-tcsa-sites panel B). Given the small number of artifacts with extreme TCSA values in the ST zone, we hesitate to conclude this is an archaeologically significant pattern of different TCSA values. The Tukey HSD pairwise differences show no pairs of eco-regions with significant differences in TCSA values (@fig-gis-tcsa-sites panel F). Overall, our results indicate no clear patterns in the distribution of TCSA values across the vegetation and eco-regional zones.

@@
